---
title: "globaltrends Public Goods Search Index (PGSI)"
subtitle: " A Google Trends-Based Measure for Public Goods Salience"
format: gfm
execute:
  echo: false
  message: false
  warning: false
---

```{r}
library(knitr)
library(nanoparquet)
library(simplecolors)
library(tidyverse)
```

The **globaltrends PGSI** is based on individual-level issue salience using data
from [Google Trends](https://trends.google.com/trends/). Our concept provides a
bottom-up complement to established measures that focus more on the quality of
infrastructure, service provision, or environmental resources. The
**globaltrends PGSI** can serve to predict demands regarding or dissatisfaction
about public goods in a country voiced by a popular majority of individuals. In
contrast to existing measures, the **globaltrends PGSI** measure captures issue
salience directly among individuals rather than media or analyst reports. As
such, it is uniquely representative of socio-political sentiment regarding
public goods, timely, and widely available. Our concept and measure offer novel
opportunities for economics, political science, and management researchers.

The current version of the **globaltrends PGSI** dataset can be accessed in
different file formats:

+ Granuar data (location-month-term)
  + [PARQUET](https://github.com/ha-pu/globaltrends_valpop/blob/main/data/valpop_pgsi.parquet)
+ Aggregated data (location-month-category)
  + [PARQUET](https://github.com/ha-pu/globaltrends_valpop/blob/main/data/valpop_pgsi_agg.parquet)
  + [XLSX](https://github.com/ha-pu/globaltrends_valpop/blob/main/data/valpop_pgsi_agg.xlsx)

The **globaltrends PGSI** is part of the
[globaltrends project](https://github.com/ha-pu/globaltrends).

+ Last update of the data: 2026-01-18
+ Coverage of the data: 2010-2025

# Index construction

## Index structure

The index uses a set of search [topics](https://web.archive.org/web/20230117193147/https://support.google.com/trends/answer/4359550/)
related to various public goods to collect data on the salience of public goods
in multiple countries. The full list of [topics](https://github.com/ha-pu/globaltrends_valpop/blob/main/input/valpop_topics.xlsx)
and [countries](https://github.com/ha-pu/globaltrends_valpop/blob/main/input/valpop_countries.xlsx)
is available in the repository. The usage of topics instead of plain-language
terms avoids issue related to translations or ambiguous meaning. Google's
classification algorithm assigns search queries to these topics. This separates
search queries for *party* in the social-gathering sense from *party* in the
sense of a political group.

Search terms are categories in five broad groups (with sub-categories):

1. Environmental public goods
2. Health-related public goods
3. Infrastructure-related public goods
4. Cultural public goods
5. Governance and civic participation

The category *Environmental public goods*, for example, consists of the
sub-categories (1) Clean air, (2) Clean water, (3) Green spaces, and (4)
Waste management. The sub-category *Clean air*, in turn, consists of the
following search topics:

+ air pollution
+ air quality index
+ smog
+ dust

In addition, the **PGSI** tracks interest in EU institutions for each country:

+ European Union
+ European Commission
+ European Parliament
+ Court of Justice of the European Union

## Search-score releveling

The Google Trends platform does not provide raw search counts, but rather
relative search volumes, which are first calculated by dividing the number of
searches for a keyword by the total searches for a given location and time, and
then normalized by the maximum search volume within each keyword batch. This
normalization, however, makes direct comparisons across different batches,
locations, or time periods difficult due to its dependency on the specific
keywords included.

To address this, we apply a releveling process to adjust all search volumes to a
common benchmark using a control keyword, ensuring that data from different
batches are comparable. We calculate comparable search scores, which divide the
search volumes of a keyword by the sum of volumes for a set of control
keywords (see [here](https://github.com/ha-pu/globaltrends_valpop/blob/main/input/valpop_topics.xlsx))
that approximate baseline search traffic. The resulting search scores are
independent of batch, location, and time, enabling fair and meaningful
comparisons of search interest across diverse keywords, countries, and
timeframes.

## Additional data

The index also provides data on related search terms for each topic-country-year
and search regions for each topic-country-year. This data offers further
insights into the context in which searches are made and where these
searches are concentrated within countries.

# Data samples

## Data structure

Ten random rows from the aggregated **PGSI** dataset for France:

```{r}
read_parquet("data/valpop_pgsi_agg.parquet") %>%
    filter(iso3 == "FRA") %>%
    slice_sample(n = 10) %>%
    kable()
```

Ten random rows from the granular **PGSI** dataset for Germany:

```{r}
read_parquet("data/valpop_pgsi.parquet") %>%
    filter(iso3 == "DEU") %>%
    slice_sample(n = 10) %>%
    kable()
```

Ten random rows from the related search-term **PGSI** dataset for Austria:

```{r}
read_parquet("data/valpop_pgsi_related.parquet") %>%
    filter(iso3 == "AUT") %>%
    slice_sample(n = 10) %>%
    kable()
```

Ten random rows from the search region **PGSI** dataset for Italy:

```{r}
read_parquet("data/valpop_pgsi_region.parquet") %>%
    filter(iso3 == "ITA") %>%
    slice_sample(n = 10) %>%
    kable()
```

## Salience of EU institutions

The plot below shows the search interest in EU institutions in the included
European countries. Seach interest is averaged across the four institutions
included and computed as an un-weighted mean across countries (black line).
For comparison, the plot also includes search scores for Germany, France,
and the UK.

```{r}
df <- read_parquet("data/valpop_pgsi_agg.parquet") %>%
    filter(category_1 == "EU Institutions") %>%
    summarise(pgsi = mean(pgsi), .by = c(iso3, year))

df %>%
    summarise(pgsi = mean(pgsi), .by = year) %>%
    mutate(iso3 = "Total") %>%
    bind_rows(filter(df, iso3 %in% c("DEU", "FRA", "GBR"))) %>%
    ggplot() +
    geom_line(aes(x = year, y = pgsi, colour = iso3)) +
    scale_color_manual(values = sc("mutedgreen3", "mutedblue3", "mutedred3", "grey5")) +
    labs(
      x = NULL,
      y = "PGSI",
      colour = NULL
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

The UK data nicely shows a peak in interest in EU institutions around the BREXIT
vote in 2016 and in 2019, when actual BREXIT was postponed multiple times (and
completed in January 2020).

## Salience of infrastructure-related public goods

The plot below shows the search interest in infrastructure-related public
goods in the included European countries. Seach interest is averaged across
all search topics included and computed as an un-weighted mean across countries
(black line). For comparison, the plot also includes search scores for the
Netherlands, Poland, and Spain.

```{r}
df <- read_parquet("data/valpop_pgsi_agg.parquet") %>%
    filter(category_1 == "Infrastructure-related public goods") %>%
    summarise(pgsi = mean(pgsi), .by = c(iso3, year))

df %>%
    summarise(pgsi = mean(pgsi), .by = year) %>%
    mutate(iso3 = "Total") %>%
    bind_rows(filter(df, iso3 %in% c("ESP", "NLD", "POL"))) %>%
    ggplot() +
    geom_line(aes(x = year, y = pgsi, colour = iso3)) +
    scale_color_manual(values = sc("mutedyellow3", "brightorange3", "mutedred3", "grey5")) +
    labs(
      x = NULL,
      y = "PGSI",
      colour = NULL
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

### Within-category variation

The plot below breaks down the search data on infrastructure-related public
goods for Spain. The aggregate show above combines search interest for six
categories:

+ Bike and pedestrian safety
+ Electriciy grid
+ Internet and telecommunication
+ Public housing
+ Public transport
+ Road infrastructure

```{r}
read_parquet("data/valpop_pgsi_agg.parquet") %>%
    filter(category_1 == "Infrastructure-related public goods" & iso3 == "ESP") %>%
    summarise(pgsi = mean(pgsi), .by = c(category_2, year)) %>%
    ggplot() +
    geom_line(aes(x = year, y = pgsi, colour = category_2)) +
    scale_color_manual(values = sc(paste0("muted", c("red", "orange", "green", "teal", "blue", "violet"), 3))) +
    labs(
      x = NULL,
      y = "PGSI",
      colour = NULL
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

In the case of the most salient topic, *public transport*, the dip in search
interest during the COVID-19 period becomes apparent.

### Topic-level variation

The plot below further breaks down the search data on public transport in Spain.
The aggregate category *public transport* shown above combines search interest
for four search topics:

+ Bus
+ Metro
+ Public transport
+ Train

```{r}
read_parquet("data/valpop_pgsi.parquet") %>%
    filter(category_2 == "Public transport" & iso3 == "ESP") %>%
    summarise(pgsi = mean(pgsi), .by = c(topic, year)) %>%
    ggplot() +
    geom_line(aes(x = year, y = pgsi, colour = topic)) +
    scale_color_manual(values = sc(paste0("muted", c("red", "orange", "green", "blue"), 3))) +
    labs(
      x = NULL,
      y = "PGSI",
      colour = NULL
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

## Salience of environmental public goods: clean water

Among the search topics related to clean water in Ireland, the topic
*drinking water quality* attrachted the greatest search interest in 2022.

```{r}
read_parquet("data/valpop_pgsi.parquet") %>%
    filter(category_2 == "Clean water" & iso3 == "IRL" & year == 2023) %>%
    summarise(pgsi = mean(pgsi), .by = topic) %>%
    ggplot() +
    geom_col(aes(x = topic, y = pgsi)) +
    labs(
      x = NULL,
      y = "PGSI"
    ) +
    theme_bw()
```

The **trending topic** feature of Google Trends allows us to further explore
the underlying search interest for this topic. For each search topic, we can get
related terms. This provides information regarding context in which a topic has
been searched. The table below shows the top 10 related search terms for
*drinking water* in Ireland in 2023.

```{r}
read_parquet("data/valpop_pgsi_related.parquet") %>%
    filter(topic == "drinking water quality" & iso3 == "IRL" & year == 2023) %>%
    select(related_topic, related_pgsi) %>%
    slice_head(n = 10) %>%
    kable()
```

## Salience of health-related public goods: emergency services

Among the search topics related to emergency services in Denmark, the topic
*ambulance* attrachted the the second highest level of search interest in 2018.

```{r}
read_parquet("data/valpop_pgsi.parquet") %>%
    filter(category_2 == "Emergency services" & iso3 == "DNK" & year == 2018) %>%
    summarise(pgsi = mean(pgsi), .by = topic) %>%
    ggplot() +
    geom_col(aes(x = topic, y = pgsi)) +
    labs(
      x = NULL,
      y = "PGSI"
    ) +
    theme_bw()
```

The **search region** feature of Google Trends allows us to further explore
the underlying search interest for this topic. For each search topic and country,
we can get the distribution of search interest across subregsion. The table below
shows the top regions regarding search for *ambulance* in Denmark in 2018.

```{r}
read_parquet("data/valpop_pgsi_region.parquet") %>%
    filter(topic == "ambulance" & iso3 == "DNK" & year == 2018) %>%
    select(region_name, region_pgsi) %>%
    arrange(-region_pgsi) %>%
    kable()
```
